---
title: "Running slidr"
author: "Sumana Srivatsa"
date: "8/2/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Creating data object

Loading all the relevant data and creating the data object with the mutation and viability scores for each type of cancer. 

```{r data_object}
library(reshape2)
library(dplyr)
library(viridis)
library(ggplot2)
library(nem)

# Set path to all data
base_folder = "~/GitHub/CrispR/Data/"
  
# Loading mutations for all cell lines
all_cancers_mut_df <- read.delim(paste0(base_folder,
                                       "MutationFiles/MutationsCCLE/CCLE_maf_allcancers.txt"),
                                stringsAsFactors = FALSE, header = TRUE)
# Filtering out silent mutations
all_cancers_mut_df <- all_cancers_mut_df %>% 
                        dplyr::filter(Variant_Classification != "Silent" )
  
# Load annotation file and meta data
cellline_annot <-  read.csv(paste0(base_folder, "ProjectDRIVE/TableS2.csv"),
                            stringsAsFactors = F)
meta_data <- read.csv(paste0(base_folder, "MutationFiles/File_metadata.csv"), 
                      header = TRUE, stringsAsFactors = FALSE)
Primary_sites <-   meta_data$Primary_site
  
# Load log2 copy number data
CN_df <- read.delim(paste0(base_folder,"/CopyNumber/CCLE_copynumber_byGene.txt"),
                     stringsAsFactors = F, header = T)
  
# Load Gistic copy number data
CN_df_gistic <- read.delim(paste0(base_folder,"/CopyNumber/CCLE_CNA_Integers.txt"),
                            stringsAsFactors = F, header=T)
CN_df_gistic <- CN_df_gistic[-which(duplicated(CN_df_gistic$Hugo_Symbol)),]

# Getting essential genes
data_rsa <- readRDS(paste0(base_folder, "ProjectDRIVE/DRIVE_RSA_data.RDS"))

# Remove the essential data from ATARiS normalised data
data_ataris <- readRDS(paste0(base_folder, "ProjectDRIVE/DRIVE_ATARiS_data.RDS"))

# Getting a list of cell lines for each cancer type
cellline_list <- lapply(Primary_sites, 
                        function(x){slidr::getCelllines(x, cellline_annot, meta_data)})
names(cellline_list) <- Primary_sites

# Getting a list of essential genes for each cancer type
essen_gene_list <- lapply(names(cellline_list), 
                        function(x){slidr::getEssentialGenes(x = x, 
                                                             celllines = cellline_list[[x]], 
                                                             data = data_rsa)})
names(essen_gene_list) <- Primary_sites

all_data = list()
fdr = 0.05
min_Nmut = 2

# Generate objects for each primary site
all_data <- lapply(Primary_sites, 
                   function(x){ slidr::prepareDataObjects(data = data_ataris, 
                                                      x = x, 
                                                      fdr = fdr, 
                                                      min_Nmut = min_Nmut, 
                                                      all_cancers_mut_df = all_cancers_mut_df, 
                                                      CN_df = CN_df_gistic, 
                                                      gistic = TRUE, 
                                                      celllines = cellline_list[[x]], 
                                                      meta_data = meta_data, 
                                                      essential_genes = essen_gene_list[[x]])})
names(all_data) <- Primary_sites

```
## Identify SL hits

Once the data is processed in the correct format, use the `identifySLHits` function to get the mutation-specific SL partner for each cancer type. 

```{r Identify_SL}
path_results <- "~/Downloads/Slidr_Results/"

hits <- lapply(all_data, 
               function(x){slidr::identifySLHits(canc_data = x, path_results = path_results)})
names(hits) <- names(all_data)

```
## Pan Cancer Analysis

Creating the data object for the pan-cancer analysis. 

```{r Pancancer}
# Choosing cell lines with both CN and viability data and the driver genes
pc_celllines  <- intersect(colnames(data_ataris),tolower(colnames(CN_df)))
driver_genes  <- unique(all_cancers_mut_df$Hugo_Symbol)
# Getting the essential genes
pc_essen_gene <- slidr::getEssentialGenes(x = NULL, data = data_rsa, celllines = pc_celllines)
# Removing the essential genes
pc_data <- list(viabilities = NULL,
                mutations = NULL,
                CNalterations = NULL,
                mutation_annot = NULL,
                primary_site = "pan_cancer")

data_ataris <- as.data.frame(t(apply(data_ataris,
                                      1,
                                      function(x){
                                           x[which(is.na(x))] <- mean(x, na.rm = TRUE)
                                           x})))
pc_data$viabilities <- data_ataris[!rownames(data_ataris) %in% pc_essen_gene,pc_celllines]

# Threshold for removing drivers with fewer than 2% mutated samples
mut_pc     <- 8 #2         
mut_thresh <- floor(mut_pc * length(pc_celllines) / 100)
CN_Thr     <- 1 # Threshold to choose only deep deletions
# Binary mutation data
mut_mat    <- slidr::prepareMutMat(x = NULL, 
                                  driver_genes = driver_genes, 
                                  samples = pc_celllines,
                                  all_cancers_mut_df = all_cancers_mut_df)

pc_data$CNalterations <- slidr::prepareCNMat(CN_df = CN_df, 
                                             samples = pc_celllines,
                                             driver_genes = driver_genes, 
                                             x = NULL)

# binarized copy number data with only deep deletions 
CN_bin <- binarize(2 - pc_data$CNalterations, threshold = CN_Thr)
CN_bin[is.na(CN_bin)] <- 0

# Updating copy numbers in mutation matrix
pc_data$mutations <- binarize((CN_bin + mut_mat), threshold = 0)

# Removing genes with < 2% mutated samples
pc_data$mutations   <- pc_data$mutations[rowSums(pc_data$mutations) >= mut_thresh, ]

# Create a mutation annotation file
mut_mat_annot <- melt(pc_data$mutations)
mut_mat_annot <- mut_mat_annot %>% dplyr::filter(value != 0)
colnames(mut_mat_annot) <- c("Hugo_Symbol","Cell_Line","Mut_Status")
# Use copy number info in the mutation annotation file
all_cancers_mut_df$Tumor_Sample_Barcode <- tolower(all_cancers_mut_df$Tumor_Sample_Barcode)
CN_alterations_df <- melt(pc_data$CNalterations)
colnames(CN_alterations_df) <- c("Hugo_Symbol","Cell_Line","CN_Value")

# Annotating mutations and adding copy number information
pc_data$mutation_annot <- left_join(mut_mat_annot,all_cancers_mut_df,
                              by = c("Hugo_Symbol" = "Hugo_Symbol","Cell_Line" = "Tumor_Sample_Barcode")) %>%
                          dplyr::select(Hugo_Symbol, Cell_Line, Mut_Status, Variant_Classification) %>%
                          dplyr::group_by(Hugo_Symbol,Cell_Line,Mut_Status) %>% 
                          dplyr::summarise(Variant_Classification = paste(Variant_Classification, collapse=";")) 

pc_data$mutation_annot <- left_join(pc_data$mutation_annot, CN_alterations_df, by = c("Hugo_Symbol","Cell_Line"))

# Replacing all the na
pc_data$mutation_annot[is.na(pc_data$mutation_annot)] <- 0

save.image("~/Downloads/Slidr_Results/pancan.Rdata")

```
Identify Synthetic lethals from pan cancer data.

```{r sl_pancancer}
# Path for results
path_results <- "~/Downloads/Slidr_Results/"

hits_pancan <- slidr::identifySLHits(canc_data = pc_data, path_results = path_results, WT_pval_thresh = 0, qval_thresh = 1)

```

Plotting p-values of the SL pairs and identifying partners that are partners for many drivers.

```{r bubble_plot_1}
hits_pancan <- read.delim("~/Downloads/Slidr_Results/Pan_cancer_8pc/Hit_List/SL_hits_pan_cancer.txt",stringsAsFactors = F)
hits_pancan$sl_partner_gene <- sapply(hits_pancan$sl_partner_gene, function(x){strsplit(x, ",")[[1]][1]})

p1 <- ggplot(hits_pancan, aes(x=driver_gene, y=sl_partner_gene, size = -log(mut_pvalue, base = 10))) + 
      geom_point(fill = "#055C95",color="#055C95", alpha=0.45) + 
      scale_size_continuous(range=c(2, 10)) +
      theme_bw() +
      xlab("Driver genes") +
      ylab("SL partner genes") +
      labs(size = "-log(p-value)")+
      theme(panel.grid.major = element_line(size = 0.1),
            panel.grid.minor = element_blank(),
            axis.ticks.x = element_blank(),
            axis.ticks.y = element_blank(),
            axis.text.y=element_text(angle=0, size=11,colour="#525252"),
            axis.text.x=element_text(angle=90, size=11,vjust=0.5 ,hjust=1, colour="#525252"),
            axis.title.x=element_text(angle=0, size=14,colour="#525252"),
            axis.title.y=element_text(angle=90, size=14,colour="#525252"),
            legend.position="bottom",
            legend.title = element_text(angle=0, size=12,colour="#525252"),
            panel.border = element_blank()) 
ggsave(filename = "~/Downloads/Slidr_Results/Pan_cancer_8pc/SLpairs_8pc_bubblePlot.pdf", p1, width = 18, height = 17, units = "in")
  
```

Plotting p-values for a given pair across different cancers.

```{r bubble_plot_2}
hits_pancan <- read.delim("~/Downloads/Slidr_Results/Pan_cancer_8pc/Hit_List/SL_hits_pan_cancer.txt",stringsAsFactors = F)

load("~/Downloads/Slidr_Results/Pan_cancer_8pc/pancan.Rdata")

# create data frame for each pair with frequency of different celllines
canc_type_df <- do.call(rbind.data.frame, lapply(1:nrow(hits_pancan), function(x){
                                                  mutations     <- pc_data$mutations
                                                  mut_celllines <- names(which(mutations[hits_pancan$driver_gene[x],] == 1))
                                                  cancer_types  <- sapply(mut_celllines, 
                                                                        function(x){paste(strsplit(x,"_")[[1]][-1], collapse = "_")})
                                                  cancer_types  <- as.data.frame(table(cancer_types))
                                                  data.frame(hits_pancan$driver_gene[x], hits_pancan$sl_partner_gene[x], cancer_types)
                                                #}
                                                
}))

colnames(canc_type_df) <- c("driver_gene", "sl_partner_gene", "canc_type", "n_samples")

min_Nmut = 2

# get the p-value for each pair and type of cancer
pval_df <- do.call(rbind.data.frame, lapply(1:nrow(canc_type_df), function(x){
                                          canc_type        <- as.character(canc_type_df$canc_type[x])
                                          driver           <- as.character(canc_type_df$driver_gene[x])
                                          sl_partner       <- as.character(canc_type_df$sl_partner_gene[x])
                                          celllines        <- grep(canc_type,colnames(pc_data$mutations))
                                          mut_samples      <- sum(pc_data$mutations[driver,celllines])
                                          if(mut_samples >= min_Nmut & (length(celllines) - mut_samples) >= min_Nmut){
                                            temp             <- pc_data
                                            temp$mutations   <- pc_data$mutations[,celllines]
                                            temp$viabilities <- pc_data$viabilities[,celllines]
                                            c(canc_type,slidr::getPval(temp,driver,sl_partner))
                                          }else
                                            c(canc_type,driver,sl_partner,1,1)
                                          
}))

colnames(pval_df)          <- c("canc_type","driver_gene", "sl_partner_gene", "WT_pvalue", "mut_pvalue")
summary_df                 <- left_join(canc_type_df, pval_df, by = c("driver_gene","sl_partner_gene","canc_type"))
summary_df$sl_partner_gene <- sapply(summary_df$sl_partner_gene, function(x){strsplit(x, ",")[[1]][1]})
summary_df                 <- cbind(paste(summary_df$driver_gene,summary_df$sl_partner_gene, sep = "_"), 
                                    summary_df,
                                    rowSums(pc_data$mutations[summary_df$driver_gene,]))
colnames(summary_df)[c(1,8)]  <- c("sl_pairs","tot_samples")
summary_df                    <- summary_df %>% dplyr::filter(mut_pvalue != 1 & WT_pvalue != 1)

# Using bubble plot to summarize the p-values in each cancer type
q1 <- ggplot(summary_df, aes(y=canc_type, x=sl_pairs, size = -log(as.numeric(as.character(mut_pvalue)), base = 10))) + 
      geom_point(fill = "#055C95",color="#055C95", alpha=0.45) + 
      scale_size_continuous(range=c(2, 8)) +
      theme_bw() +
      ylab("Primary sites") +
      xlab("Driver and SL partner genes") +
      labs(size = "-log(p-value)")+
      theme(panel.grid.major = element_line(size = 0.1),
            panel.grid.minor = element_blank(),
            axis.ticks.x = element_blank(),
            axis.ticks.y = element_blank(),
            axis.text.x = element_text(angle = 90, vjust=1, hjust=1, size=11,colour="#525252"),
            axis.text.y = element_text(angle = 0, size=11, vjust=1, hjust=1, colour="#525252"),
            axis.title.x=element_text(angle = 0, size=14,colour="#525252"),
            axis.title.y=element_text(angle=90, size=14,colour="#525252"),
            legend.position="bottom",
            legend.title = element_text(angle=0, size=12,colour="#525252"),
            panel.border = element_blank()) +
    scale_y_discrete(labels = c("Autonomic ganglia", "Bone", "Breast", "CNS", "Endometrium", "Blood", "Kidney", "Large intestine", "Liver", "Lung", "Oesophagus", "Ovary", "Pancreas", "Pleura", "Skin", "Soft tissue", "Stomach", "UADT", "Urinary tract") )

# Using a heatmap to describe the p-values in each cancer type
q2 <- ggplot(summary_df, aes(sl_pairs,canc_type)) +
        geom_tile(aes(fill = -log(as.numeric(as.character(mut_pvalue))))) +
        scale_fill_viridis(option="D", begin = 0.1, end = 0.9, alpha = 0.3) +
        theme_bw() +
        ylab("Primary sites") +
        xlab("Driver and SL partner genes") +
        labs(fill = "-log(p-value)")+
        theme(panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            #axis.ticks.x = element_blank(),
            #axis.ticks.y = element_blank(),
            axis.text.x = element_text(angle = 90, vjust=1, hjust=1, size=11,colour="#525252"),
            axis.text.y = element_text(angle = 0, size=11, vjust=1, hjust=1, colour="#525252"),
            axis.title.x=element_text(angle = 0, size=14,colour="#525252"),
            axis.title.y=element_text(angle=90, size=14,colour="#525252"),
            legend.position="bottom",
            legend.title = element_text(angle=0, size=12,colour="#525252"))+
            #panel.border = element_blank()) +
        scale_y_discrete(labels = c("Autonomic ganglia", "Bone", "Breast", "CNS", "Endometrium", 
                                    "Blood", "Kidney", "Large intestine", "Liver", "Lung", "Oesophagus", 
                                    "Ovary", "Pancreas", "Pleura", "Skin", "Soft tissue", "Stomach", "UADT", "Urinary tract") )


ggsave(filename = "~/Downloads/Slidr_Results/Pan_cancer_8pc/SLpairs_canctype_8pc_bubblePlot.pdf", q1, width = 35, height = 9, units = "in")
ggsave(filename = "~/Downloads/Slidr_Results/Pan_cancer_8pc/SLpairs_canctype_8pc_heatmap.pdf", q2, width = 35, height = 9, units = "in")
```
Plotting stacked bar plot for number of samples.

```{r stacked_barplot}
barplot_df <- summary_df %>% dplyr::distinct(driver_gene, n_samples, canc_type)
barplot_df$canc_type <- factor(barplot_df$canc_type, 
                               labels = c("Autonomic ganglia", "Bone", "Breast", "CNS", "Endometrium", "Blood", "Kidney", "Large intestine", "Liver", "Lung", "Oesophagus",
"Ovary", "Pancreas", "Pleura", "Skin", "Soft tissue", "Stomach", "UADT", "Urinary tract"))

coloursSites <- c("#CC99BB", "#AA4488", "#771155", "#77AADD",
"#4477AA", "#114477", "#77CCCC", "#44AAAA", "#117777", "#88CCAA",
"#44AA77", "#117744", "#DDDD77", "#AAAA44", "#777711", "#DDAA77",
"#AA7744", "#774411", "#DD7788")

p2 <- ggplot(barplot_df, aes(fill = canc_type, x = driver_gene, y = n_samples)) +
        geom_bar(stat = "identity") +
        theme_bw() +
        ylab("Number of mutated cell lines") +
        xlab("Driver genes") +
        labs(fill = "Primary site")+
        theme(panel.grid.major = element_blank(),
             panel.grid.minor = element_blank(),
             panel.border = element_blank(),
             axis.line = element_line(colour = "#525252"),
             axis.text.x = element_text(angle = 90, vjust=0.4, hjust=1, size=11,colour="#525252"),
             axis.text.y = element_text(angle = 0, size=11, vjust=1, hjust=1, colour="#525252"),
             axis.title.x=element_text(angle = 0, size=14,colour="#525252"),
             axis.title.y=element_text(angle=90, size=14,colour="#525252"),
             legend.position="right",
             legend.title = element_text(angle=0, size=12,colour="#525252"))+
        scale_fill_manual(values = coloursSites)+
        scale_y_continuous(breaks = seq(0, 280, by = 40))

ggsave("~/Downloads/Slidr_Results/Pan_cancer_8pc/drivers_canctype_8pc_barplot.pdf", p2, width = 17, height = 8, units = "in")
```
Group the SL partners and learn the subset structures between mutated genes. We use the `nem` package for this.

```{r subset_relations}
grouped_hits <- hits_pancan %>% 
                dplyr::group_by(sl_partner_gene) %>% 
                summarise(drivers = paste(driver_gene, collapse = ","))

for(i in 1:nrow(grouped_hits)){
  temp_drivers <- sapply(grouped_hits$drivers[i], function(x){strsplit(x,",")[[1]]})
  if(length(temp_drivers) >= 2){ # running for two or more genes
    D            <- t(pc_data$mutations[temp_drivers,])
    control      <- set.default.parameters(unique(colnames(D)),type="mLL",pcombi = TRUE, trans.close=TRUE, para = c(1e-4,1e-4))
    temp_res     <- nem(D, inference = "ModuleNetwork", control=control, verbose=FALSE)
    temp_adj_mat <- transitive.reduction(as(temp_res$graph, "matrix"))
  
    pdf(file =  paste(path_results, "Subset_plots/", grouped_hits$sl_partner_gene[i], ".pdf", sep = ""))
    plot(as(temp_adj_mat, "graphNEL"),
       attrs=list(node=list(shape = "ellipse", 
                            color="olivedrab3", 
                            fontcolor="deepskyblue4",
                            fontsize=12),
                  edge=list(color="gray56", 
                            arrowhead = "normal")),
                  main= grouped_hits$sl_partner_gene[i], col.main = "gray56")
    dev.off()
  }
  
}

```
